---
title: "Multinomial Regression"
output: html_notebook
---

```{r}
library(stargazer)
library(nnet)
library(caret)  

```

```{r}
df <- read.csv("~/Documents/GitHub/CSS206/ReplicationData/cleaned.csv")

dim(df)
#originial data frame has 7850 obs X 123 variables

str(df)

```

Variables of interest :

-   Dependent variable - voteintentionspain (categorical - order doesn't matter)

    -   1 PSOE

    -   2- PP

    -   3- Podemos

    -   4 - Ciduadanos

    -   7- ERC (Catalonia)

    -   23 - Vox

-   Independent variables

    -   education - as.factor(edu3) - this variable represents 1 "Lower secondary", 2 "Upper secondary", 3 "Tertiary"

    -   income - dhincome_all - Normalized income (range 0–1)

    -   female - (0,1)

    -   nativism (percent)

    -   msexism (percent)

    -   femdemonstrate (1- YES 2- No)

-   Why we chose these variables?

    -   nativism- because anti-immigrant rhetoric is used in political campaigns to varying degrees across political spectrum

    -   do people who protest vote for feminist parties? - feminist protest

    -   msexism - are people who are more sexist, voting for parties with these beliefs

```{r}
#CLEAN the data 

#POLITICAL PARTIES OF INTEREST 

# Step 1: Define the parties we want to look at
keep_codes <- c(1, 2, 3, 4, 7, 23)

# Step 2: Filter the DataFrame
df_clean <- df[df$voteintentionspain %in% keep_codes, ]


#IND VARS OF INTEREST 

# Step 1: Select only the relevant columns
df_clean <- df_clean[, c(
  "voteintentionspain",   # Dependent variable (now a factor with 6 categories)
  "edu3",                # For education
  "dhincome_all",         # Income
  "female",               # 0 = male, 1 = female
  "nativism",             # Percent
  "msexism",              # Percent
  "femdemonstrate"        # 1 = Yes, 2 = No
)]

# Recode: 2 (No) becomes 0
df_clean$femdemonstrate[df_clean$femdemonstrate == 2] <- 0

table(df_clean$voteintentionspain)

```

```{r}
library(ggplot2)
library(dplyr)

# Create a proportion table as a data frame
df_clean %>%
  count(voteintentionspain) %>%
  mutate(percentage = n / sum(n) * 100) %>%
  ggplot(aes(x = reorder(voteintentionspain, -percentage), y = percentage)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = sprintf("%.1f%%", percentage)), vjust = -0.5) +
  labs(title = "Vote Intention in Spain",
       x = "Vote Intention",
       y = "Percentage") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#Removing NA's
str(df_clean)
df_clean <- df_clean[!is.na(df_clean$femdemonstrate) & !is.na(df_clean$dhincome_all), ]

col_missing_values <- sapply(df_clean, function(x) sum(is.na(x)))

# Print the number of missing values for each column
print(col_missing_values)

#3034 observations after NA's removed 

write.csv(df_clean, "df_clean.csv", row.names = FALSE)

```

#Model 1

Got rid of education from the model

```{r}
#Multinomial Regression 
library ( nnet )

#Model 1 - Basic multinomial model 
mnl.fit <- multinom(voteintentionspain ~ dhincome_all + female + nativism + msexism + femdemonstrate , Hess=T, model=T, data=df_clean, maxit=200) #model

summary(mnl.fit)

```

```{r}

#Create train/test split
set.seed(123) 
train_idx <- createDataPartition(df_clean$voteintentionspain, p = 0.8, list = FALSE)
train_data <- df_clean[train_idx, ]
test_data <- df_clean[-train_idx, ]

# Fit model on training data
mod1 <- multinom(voteintentionspain ~ dhincome_all + 
                    nativism + msexism + femdemonstrate,
                    Hess = TRUE, model = TRUE, data = train_data, maxit = 200)

# Step 4: Predict on test data
pred_class <- predict(mnl.fit, newdata = test_data)

# Step 5: Evaluate accuracy
confusion_matrix <- table(Predicted = pred_class, Actual = test_data$voteintentionspain)
accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)

print(confusion_matrix)
print(paste("Accuracy:", round(accuracy, 3)))


#This model is predicting 1 (PSOE) a lot more than the other categories. Why might this be? 
#Accuracy 37.9 % 
```

```{r}
library(caret)

df_clean$voteintentionspain <- as.factor(df_clean$voteintentionspain)


# Define training control with 5-fold CV
train_control <- trainControl(method = "cv", number = 10)

# Train the multinomial model using caret's train()

cv_model <- train(voteintentionspain ~ dhincome_all + 
                    nativism + msexism + femdemonstrate,
  data = df_clean,
  method = "multinom",
  trControl = train_control,
  MaxNWts = 10000,
  trace = FALSE
)

# View results
print(cv_model)

```

We chose to downsample the 1's PSOE because they made up 1230/ 3034 40% of observations, creating class imbalance.

```{r}
# Confusion Matrix
confusion_matrix <- table(Predicted = pred_class, Actual = test_data$voteintentionspain)

# Calculate accuracy
accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)

print(confusion_matrix)
print(paste("Accuracy:", round(accuracy, 3)))

#accuracy very slightly improved performance 37.7% 
```

```{r}
#RANDOM FOREST 
library(randomForest)

# Make sure voteintentionspain is a factor
train_data$voteintentionspain <- as.factor(train_data$voteintentionspain)

# Fit the model
rf_model <- randomForest(voteintentionspain ~ dhincome_all + female + 
                           nativism + msexism + femdemonstrate,
                         data = train_data,
                         ntree = 500,           # Number of trees
                         mtry = 3,              # Number of variables tried at each split (default = sqrt(p))
                         importance = TRUE)     # Enables variable importance measures

```

```{r}
# Predict on test set
preds <- predict(rf_model, newdata = test_data)

# Create confusion matrix
conf_mat <- table(Predicted = preds, Actual = test_data$voteintentionspain)

# Calculate accuracy
accuracy <- sum(diag(conf_mat)) / sum(conf_mat)
print(paste("Accuracy:", round(accuracy * 100, 2), "%"))

importance(rf_model)
varImpPlot(rf_model)

# "Accuracy: 35.37 %"
```

-   `msexism`, `dhincome_all`, and `nativism` are your **top predictors** of voting intention.

-   Gender (`female`) and participation in women's rights demonstrations (`femdemonstrate`) are **less predictive**.
